% Division of Planetary Sciences electronic abstract form.

\documentstyle[11pt,dpsab]{article}
\pagestyle{empty}

\runningident{}
\session{0}{00}
\subdate{June 1, 1997}
\societydiv{Division for Planetary Sciences}

\presentationtype{oral}     % REQUIRED

\firstname{Britt R.}            % REQUIRED
\lastname{Scharringhausen}             % REQUIRED

\sponsor{Philip D. Nicholson}             % REQUIRED FOR NONMEMBERS


\firstaddress{518 Space Sciences}        
\secondaddress{Cornell University}
\thirdaddress{}
\fourthaddress{}
\city{Ithaca}
\state{NY}
\zip{14850}
\country{USA}

\authorphone{607-255-6472}          % REQUIRED
       
\authorfax{607-255-5907}            % REQUIRED

\authoremail{brs@astrosun.tn.cornell.edu}          % REQUIRED

\membertype{STUDENTNON}           % REQUIRED

\firstpresent{YES}         % REQUIRED FOR NONMEMBERS

\begin{document}

\category{13}         % REQUIRED

\title{Ground-Based Observations of the Saturn Ring-Plane Crossings
of August and November 1995}     % REQUIRED

\author{B. R. Scharringhausen, P. D. Nicholson}
\affil{Cornell}

\author{K. Matthews}
\affil{Caltech}

\author{P. McGregor}
\affil{MSSSO}

\author{R. G. French}
\affil{Wellesley}

\author{L. Lim}
\affil{MIT}


\begin{abstract}

We present two sets of ground-based observations of the 1995 Saturn ring plane
crossings.  Observations on the 5-meter Hale Telescope at Palomar were made 
on August 13 - 17 (UT) and November 21, 22, and 23.  A 2.276~$\mu$m filter 
with a bandpass of 0.17~$\mu$m was used.  This is in a methane absorption 
band where the albedo of Saturn is low, reducing scattered light from the 
planet.  The $256\times 256$ InSb Cassegrain Infrared Camera has a plate scale
of $0.125''$ per pixel.  The August observations followed the Earth crossing
on August 10, and the November observations were at the end of the November 
17-21 Sun crossing with a ring opening angle of $2.7^\circ$.  Additional 
observations were made on August 9, 10, and 11 (UT) with the CASPIR 
camera on the ANU 2.3 meter telescope at Siding Spring Observatory (SSO), 
Australia.  A 2.34~$\mu$m narrow band filter was used, and the plate scale 
was $0.25''$ per pixel.  The August 10 SSO observations run up to 21:00 
UT, just before the ring crossing.  

The principal goal of the Palomar observations was to acquire accurate
orbital data on both the ring moons and the classical satellites by observing 
their eclipses and occultations by Saturn, as well as several mutual events 
between larger satellites.  This program was very successful, resulting in
good data for a total of 20 events involving Mimas, Enceladus, Tethys, Dione,
Rhea, Titan, and Hyperion, as well as 4 eclipses of Janus and Epimetheus.
[Nicholson, {\it et al.} {\sl BAAS} {\bf 23,} 1073 (1996)]  We present here
radial profiles of ring brightness at 2.3~$\mu$m, their temporal evolution
through the August ring plane crossing, and a search for any east-west 
asymmetries such as are seen in the HST data [Nicholson, {\it et al.} 
{\sl Science} {\bf 272,} 509 (1996)].  The Palomar observations on November
21 were made under conditions of excellent ($0.3''$) seeing, and provide
additional astrometric data for Janus, Epimetheus, Pandora and Prometheus,
as well as high-quality profiles of brightness for the unlit side of the rings.

\end{abstract}

\admininfo
\end{document}
