% HST observations of small saturnian satellites and F ring clumps
% Colleen A. McGhee, Philip D. Nicholson, Cornell Univ., Richard G. 
% French, Wellesley College

\def\deg{$^{\circ}$}
% DPSFORM.TEX -- DPS meeting abstract electronic form.

% DON'T remove the following comment; it identifies the form.
%
% Division of Planetary Sciences electronic abstract form.

% The following two lines are REQUIRED.  Leave them alone.

\documentstyle[11pt,dpsab]{article}
\pagestyle{empty}

\runningident{}
\session{0}{00}
\subdate{}
\societydiv{Division for Planetary Sciences}

\presentationtype{oral}     % REQUIRED

\newphd{no}               % OPTIONAL

\pressabstract{}        % OPTIONAL

\firstname{Colleen A.}            % REQUIRED
\lastname{McGhee}             % REQUIRED

\firstaddress{408 Space Sciences Building}        
\secondaddress{Cornell University}
\thirdaddress{}
\fourthaddress{}
\city{Ithaca}
\state{NY}
\zip{14850}
\country{USA}

\authorphone{(607) 255-6307}          % REQUIRED
       
\authorfax{(607) 255-5907}            % REQUIRED

\authoremail{mcghee@astrosun.tn.cornell.edu}          % REQUIRED

\instructions{Category:  Saturn RPX satellites/ring arcs}        % OPTIONAL

\sessionchair{no}        % OPTIONAL

\membertype{STUDENTDPS}           % REQUIRED

\firstpresent{no}         % REQUIRED FOR NONMEMBERS

\begin{document}

%    22. Other (describe in "\instructions" field above)
% Type in your ONE category code (the number only) between the braces 
% next to the word "\category", e.g. "\category{5}".

\category{22}            % REQUIRED


\title{HST observations of small saturnian satellites and F ring clumps}               % REQUIRED

\author{C. A. McGhee, P. D. Nicholson}
\affil{Cornell Univ.}
\author{R. G. French}
\affil{Wellesley College}

\begin{abstract}

Hubble Space Telescope (HST) observations of Saturn's ring and satellite
system during the August 1995 Earth ring plane crossing and
during the November 1995 solar crossing are presented. Very low 
ring brightnesses during these periods allowed 
detections of faint satellites Pandora and Prometheus and
lagrangian satellites Telesto, Calypso and Helene, and
identification of objects
determined to be in or near the F ring (Nicholson {\it et. al} [1996], 
{\it Science} {\bf 272}, 509).
In August, three objects identified as 1995S5-S7 were seen in multiple 
HST orbits, while November's data show
at least two elongated "clumps."   Attempts to correlate these objects
with those found in the May Earth-crossing HST data of Bosh and
Rivkin ([1996], {\it Science} {\bf 272}, 518) (1995S1-S4) have resulted 
in the possible linking of
1995S3 with 1995S6, if both objects are assumed to be exactly within the
F ring.  While 1995S1 is not seen convincingly
in subsequent data sets, this object may be present in some three August
HST images and two November images.  
Data taken with
the Canada-France-Hawaii Telescope (CFHT) with an adaptive optics
system during the August crossing resulted in the identification of 
several new objects (Roddier et al [1996] IAUC 6407); 
object 1995S9 has been recovered in five August HST images.

The puzzling question of Prometheus' 18.8\deg lag has spurred
the development of new analysis and measurement techniques for better
determination of orbits of these bodies, and for
the first time these techniques are applied consistently to all 
three HST data sets. 
The new techniques include subtraction of templates from the images, as 
opposed
to the ring-pair subtraction used formerly to identify and measure the new
objects, and 
the inclusion of orbital smear in modelling the image shapes, 
rather than using a simple centroid
procedure to measure the positions.  The template subtraction has
several advantages, including a better ability to distinguish among
the many small objects that are very close together in some of the
frames. Our goals are to fit the remaining objects that were
seen in the HST data, as well as to link the
objects from all the HST data sets together with ground-based 
observations.
\end{abstract}

% The following two lines are REQUIRED.  Leave them alone.

\admininfo
\end{document}
